Methods for understanding the variable importance of local explanations of black-box models

### 2021-11-05 _paper.rmd
init fifa and penguins structure for coffee and chocolates.

# Case Studies {#sec:casestudies}
## 1) Penguins species classification
## 2) FIFA wage regression
## 3)?
## 4)?

>>>>

# Case Studies {#sec:casestudies}
## 1) penguins, species classification

```{r casepenguins, echo=F, out.width = "100%", fig.cap = "Penguins classificiation TODO: XXX"}
knitr::include_graphics("./figures/case_penguins.pdf")
```

## 2) Chocolates, milk/dark chocolate classification

TODO:XXX

## 3) FIFA, wage regression
```{r casefifa, echo=F, out.width = "100%", fig.cap = "FIFA 2020, regressing wages [log 2020 Euros] from skill measures. TODO: XXX"}
knitr::include_graphics("./figures/case_fifa.pdf")
```

## 4) Coffee, rating regression

TODO:XXX

regress score color on species?

### 2021-11-05 _paper.tex
init fifa and penguins structure for coffee and chocolates.

Another way of thinking about this would be: what would the predictions be if the explanation were different. This is similar to the idea of \emph{ceteris paribus} profiles (\textbf{biecek\_ceterisparibus\_2020?}). \emph{Ceteris paribus} is Latin for ``other things held constant'' or ``all else unchanged.'' The profiles visualize what `what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant. In contrast to \emph{ceteris paribus} profiles, touring methods visualize all observations rather than one, and variables are not treated as independent, that is the projection has an orthonormal basis; when one variable is rotated out-of-frame, other variables are effectively rotated into-the-frame.

@@

The benefit of having dynamic interaction with data is predicated on a reasonably small render time. It is important to preprocess as much work as possible so that application resources can be used efficiently. Below we discuss the steps and details of the reprocessing.

@@

The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters. The application is made with \textbf{shiny} (Chang et al. 2021). The global view is created first with \textbf{ggplot2} (Wickham 2016) which is then rendered to an interactive html widget with \textbf{plotly} (Sievert 2020). The tour view is created in \textbf{spinifex} (Spyrison and Cook 2020), which creates the manual tour basis array, and facilitates similar rendering to html widget.

@@

\hypertarget{extend-spinifex-consume-dalex-treeshap}{%
\subsection{Extend spinifex, consume DALEX \& treeshap}\label{extend-spinifex-consume-dalex-treeshap}}

\hypertarget{preprocess}{%
\subsection{Preprocess}\label{preprocess}}

\hypertarget{runtime-rendering}{%
\subsection{Runtime rendering}\label{runtime-rendering}}

@@

\hypertarget{penguins-species-classification}{%
\subsection{1) Penguins species classification}\label{penguins-species-classification}}

\hypertarget{fifa-wage-regression}{%
\subsection{2) FIFA wage regression}\label{fifa-wage-regression}}

\hypertarget{section}{%
\subsection{3)?}\label{section}}

\hypertarget{section-1}{%
\subsection{4)?}\label{section-1}}

@@

@@

>>>>

Another way of thinking about this would be: what would the predictions be if the explanation were different. This is similar to the idea of \emph{ceteris paribus} profiles (Biecek 2020). \emph{Ceteris paribus} is Latin for ``other things held constant'' or ``all else unchanged.'' The profiles visualize what `what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant. In contrast to \emph{ceteris paribus} profiles, touring methods visualize all observations rather than one, and variables are not treated as independent, that is the projection has an orthonormal basis; when one variable is rotated out-of-frame, other variables are effectively rotated into-the-frame.

@@

The benefit of having dynamic interaction with data is predicated on a reasonably small render time. It is important to preprocess as much work as possible so that application resources can be used efficiently. The work remaining at runtime should be solely responding to inputs and the rendering of figures and tables. Below we discuss the steps and details of the reprocessing.

@@

@@

The above described method and application are implemented as an open source \textbf{R} package, \textbf{cheem} \emph{TODO:XXX site github? cran?}. Preprocessing was facilited with models created via \textbf{randomForest} {[}liaw\_classification\_2002{]}, and explanations calculated with \textbf{treeshap} (Kominsarczyk et al. 2021).
The application is made with \textbf{shiny} (Chang et al. 2021). The tour visual is an extension of \textbf{spinifex} (Spyrison and Cook 2020). Both views are created first with first with \textbf{ggplot2} (Wickham 2016) and then rendered as interactive html widget with \textbf{plotly} (Sievert 2020). The free ebook, \emph{Explanatory Model Analysis} (Biecek and Burzykowski 2021) was a huge boon to understanding local explanations and how to apply them.

\hypertarget{preprocessing-1}{%
\subsection{Preprocessing}\label{preprocessing-1}}

Provided with tidy and transformed data, the function \texttt{cheem\_ls} is a wrapper function that facilitates the creation of random forest models and calculates the tree SHAP for all observations. It returns an a list containing data frames to be consumed by the application, higher level model performance, and runtime information. The return of this function can be then be saved to an .rds file for consumption in the application.

\hypertarget{linking-brushing-and-highlighting}{%
\subsection{Linking brushing and highlighting}\label{linking-brushing-and-highlighting}}

The dynamic interaction with the global view is critical to the selection of the primary and comparison observations. Linked brushing also for a rectangular selection of observations. These points are highlighted within the other space of the global view \emph{and} crucially available for downstream consumption, for instance highlighting in the tour. The distinction between the former and latter is subtle, but important for developers to grasp. In the former, self-contained html widgets can highlight directly within javascript without evaluation evaluation downstream in other reactive functions. In the latter, the use of \texttt{plotly::event\_data()} returns the identity of the selected observations. This will require a reactive flush, but after that, the sky is the ceiling.

@@

\hypertarget{penguins-species-classification}{%
\subsection{1) penguins, species classification}\label{penguins-species-classification}}

{\centering \includegraphics[width=1\linewidth]{./figures/case_penguins}

\caption{Penguins classificiation TODO: XXX}\label{fig:casepenguins}
\end{figure}

\hypertarget{chocolates-milkdark-chocolate-classification}{%
\subsection{2) Chocolates, milk/dark chocolate classification}\label{chocolates-milkdark-chocolate-classification}}

TODO:XXX

\hypertarget{fifa-wage-regression}{%
\subsection{3) FIFA, wage regression}\label{fifa-wage-regression}}

{\centering \includegraphics[width=1\linewidth]{./figures/case_fifa}

\caption{FIFA 2020, regressing wages [log 2020 Euros] from skill measures. TODO: XXX}\label{fig:casefifa}
\end{figure}

\hypertarget{coffee-rating-regression}{%
\subsection{4) Coffee, rating regression}\label{coffee-rating-regression}}

TODO:XXX

regress score color on species?

@@

We would like to thank Professor Przemyslaw Biecek for his input in early in the project and to the broader MI\^{}2 lab group for the \textbf{DALEX} ecosystem of \textbf{R} and \textbf{Pyhton} packages. This research was supported in part by Australian government Research Training Program (RTP) scholarships.

@@

\leavevmode\hypertarget{ref-biecek_ceterisparibus_2020}{}%
Biecek, Przemyslaw. 2020. \emph{{ceterisParibus}: {Ceteris} {Paribus} {Profiles}}. \url{https://CRAN.R-project.org/package=ceterisParibus}.

\leavevmode\hypertarget{ref-biecek_explanatory_2021}{}%
Biecek, Przemyslaw, and Tomasz Burzykowski. 2021. \emph{Explanatory {Model} {Analysis}: {Explore}, {Explain}, and {Examine} {Predictive} {Models}}. CRC Press.
